---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
#install.packages("nflfastR")
```

```{r}
misery <- function(wins, total_yards){
  misery <- vector()
  for (i in 1:length(wins)) {
    
    if(total_yards[i] > 6653){
      misery[i] = 0
    }else{
      misery[i] = 3 - ((total_yards[i]-4000)/2653)^4 * 3
    }
    
    if(wins[i] > 16){
      misery[i] = misery[i] 
    }else{
      misery[i] = misery[i] + (7 - (wins[i] / 16) * 7)
    }
  }

  signif(misery, 3)
}
```
This function creates a misery index for a team's season based off the amount of wins that the 
team has and the yards that they compiled on offense. It is weighted more towards wins since that is 
what gets fanse excited.

```{r}
library(nflfastR)
library(dplyr)
library(ggplot2)
library(ggrepel)
pbp20 <- load_pbp(2020)
#names(pbp20) 

pbp20home <- pbp20 %>%
  select(season, week, home_team, result) %>%
  rename(team = home_team) %>% 
  distinct(week, team, .keep_all = TRUE)

pbp20away <- pbp20 %>%
  select(season, week, away_team, result) %>%
  rename(team = away_team) %>% 
  mutate(result = -result) %>% 
  distinct(week, team,.keep_all = TRUE)

pbp20wins <- rbind(pbp20away,pbp20home) %>% 
  mutate(win = case_when(
    result > 0  ~ 1,
    result < 0  ~ 0,
    result == 0 ~ .5
  )) %>% 
  group_by(team) %>% 
  summarise(wins = sum(win))
  

stats20 <- pbp20 %>% 
  filter(!is.na(posteam)) %>% 
  group_by(posteam) %>% 
  rename(team = posteam) %>% 
  summarise(passing_yards = sum(passing_yards, na.rm = TRUE),
            rushing_yards = sum(rushing_yards, na.rm = TRUE)) 

offense <- left_join(pbp20wins, stats20 )
offense
offense <- offense %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr')) %>% 
  mutate(misery = misery(wins, passing_yards+rushing_yards))

offense %>%
  ggplot(aes(x = rushing_yards, y = passing_yards)) +
  #horizontal line with mean passing yards
  geom_hline(yintercept = mean(offense$passing_yards), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean rushing yards
  geom_vline(xintercept =  mean(offense$rushing_yards), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the teams with the right colors
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = offense$team_color, cex= offense$wins, alpha = .6)+
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=paste(team, " Misery: ", misery))) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Offensive Rushing Yards for 2020",
       y = "Offensive Passing Yards for 2020",
       title = "Offense Impact on Wins for 2020",
       caption = "Data: @nflfastR") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) +
  #make ticks look nice
  #if this doesn't work, `install.packages('scales')`
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
```

This plot shows us the number of passing yards and the number of rushing yards for a team in the 2020
season. The size of the point on the plot corresponds to the number of wins that the team had in 2020. 
The color of the point corresponds to the team's color. 

```{r}
pbp <- load_pbp(1999:2020)
#names(pbp20) 

pbphome <- pbp %>%
  select(season, week, home_team, result) %>%
  rename(team = home_team) %>% 
  distinct(week, team, season, .keep_all = TRUE)

pbpaway <- pbp %>%
  select(season, week, away_team, result) %>%
  rename(team = away_team) %>% 
  mutate(result = -result) %>% 
  distinct(week, team, season, .keep_all = TRUE)

pbpwins <- rbind(pbpaway,pbphome) %>% 
  mutate(win = case_when(
    result > 0  ~ 1,
    result < 0  ~ 0,
    result == 0 ~ .5
  )) %>% 
  group_by(team, season) %>% 
  summarise(wins = sum(win))


stats <- pbp %>% 
  filter(!is.na(posteam)) %>% 
  group_by(posteam, season) %>% 
  rename(team = posteam) %>% 
  summarise(passing_yards = sum(passing_yards, na.rm = TRUE),
            rushing_yards = sum(rushing_yards, na.rm = TRUE)) 

offense <- left_join(pbpwins, stats )
offense <- offense %>%
  left_join(teams_colors_logos, by = c('team' = 'team_abbr'))

jaxoffense <- offense %>% 
  filter(team == 'JAX') %>% 
  mutate(misery = misery(wins, rushing_yards + passing_yards))


jaxoffense %>%
  ggplot(aes(x = rushing_yards, y = passing_yards)) +
  #horizontal line with mean passing yards
  geom_hline(yintercept = mean(jaxoffense$passing_yards), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean rushing yards
  geom_vline(xintercept =  mean(jaxoffense$rushing_yards), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the teams with the right colors
  #cex controls point size and alpha the transparency (alpha = 1 is normal)
  geom_point(color = "#008080",cex= jaxoffense$wins, alpha = .6)+
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=paste0(season," Misery: ",misery))) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "Offensive Rushing Yards",
       y = "Offensive Passing Yards",
       title = "JAX Offense Impact on Wins 1999:2020",
       caption = "Data: @nflfastR") +
  #uses the black and white ggplot theme
  theme_bw() +
  #center title with hjust = 0.5
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold")
  ) +
  #make ticks look nice
  #if this doesn't work, `install.packages('scales')`
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) 
```


```{r}
qbr_all <- load_pbp(1999:2020)

qbr_all_filter <- qbr_all %>%
  select(season, week, home_team, away_team, posteam, play_type, result, season_type, third_down_converted, third_down_failed,
         incomplete_pass, interception, pass_attempt, pass_touchdown, complete_pass,passer_player_name,
         passing_yards, pass, result, sack
         ) %>%
  mutate(result = case_when(
    away_team == posteam  ~ -result,
    home_team == posteam  ~ result)
    ) %>%
  filter(play_type %in% c("pass"), season_type == "REG") %>%
  group_by(season, week, posteam) %>%
  summarise(
    pass_plays = n(),
    sacks = sum(sack, na.rm = TRUE),
    passes_attempted = sum(pass_attempt, na.rm = TRUE) - sacks,
    incompletions = sum(incomplete_pass, na.rm = TRUE),
    completions = passes_attempted - incompletions,
    pass_touchdowns = sum(pass_touchdown, na.rm = TRUE),
    passing_yards = sum(passing_yards, na.rm = TRUE),
    interceptions = sum(interception, na.rm = TRUE),
    third_down_converted = sum(third_down_converted, na.rm = TRUE),
    third_down_failed = sum(third_down_failed, na.rm = TRUE),
    third_down_conversion_rate = signif(third_down_converted/(third_down_failed + third_down_converted), 3),
    result = unique(result)
    ) %>%
  mutate(win = case_when(
    result > 0  ~ 1,
    result < 0  ~ 0,
    result == 0 ~ .5
    )) %>%
  mutate(a = ((completions/passes_attempted - .3) * 5),
         b = ((passing_yards/passes_attempted - 3) * .25),
         c = ((pass_touchdowns/passes_attempted) * 20),
         d = (2.375 - (interceptions/passes_attempted * 25))
         ) %>%
  mutate(
    a = case_when(
    a >= 2.375 ~ 2.375,
    a >= 0 ~ a,
    a < 0 ~ 0
    ), b = case_when(
    b >= 2.375 ~ 2.375,
    b >= 0 ~ b,
    b < 0 ~ 0
    ), c = case_when(
    c >= 2.375 ~ 2.375,
    c >= 0 ~ c,
    c < 0 ~ 0
    ), d = case_when(
    d >= 2.375 ~ 2.375,
    d >= 0 ~ d,
    d < 0 ~ 0
    ),
    qbr = signif((a + b + c +d)/6 * 100,4),
    win = as.logical(win),
    season = as.factor(season)
    ) %>%
  select(-c(a,b,c,d)) %>%
  filter(!is.na(win))

ggplot(qbr_all_filter, aes(x=season, y= qbr, fill=win)) +
  geom_boxplot(position = position_dodge(0.9) ) +
  labs(title = "QBR For Winning and Losing",
       x = "Season",
       y = "QBR") +
  scale_fill_discrete(name = "Result", labels = c("Lose", "Win")) +
  theme_minimal() +
  geom_hline(yintercept= mean(qbr_all_filter$qbr),color = "black", linetype = "dashed", alpha=0.5) +
  geom_hline(yintercept = mean(qbr_all_filter$qbr[qbr_all_filter$win == TRUE]), color = "#00BFC4", size = 1, linetype = "dashed") +
  geom_hline(yintercept = mean(qbr_all_filter$qbr[qbr_all_filter$win == FALSE]), color = "#F8766D", size = 1, linetype = "dashed")

```

